---
title: "Chapter 4"
format: 
  html:
    code-fold: false
    code-tools: true
editor: source
author: Lilla Gurtner
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = F, message = F, error=F, include = F}

# knitr global options ----
knitr::opts_chunk$set(fig.pos = 'H',
                      echo = T,
                      message = F,
                      warning = F, 
                      dpi = 600,
                      fig.align = "center", 
                      fig.asp = 0.62) # golden ratio

library(tidyverse)
library(tidybayes) # nice ploting
library(brms)
library(wesanderson)

theme_set(theme_tidybayes())

```

# 4E1. 

In the model definition below, which line is the likelihood? 

*yi ∼ Normal(μ, σ)*

μ ∼ Normal(0, 10) 

σ ∼ Exponential(1)

# 4E2. 

In the model definition just above, how many parameters are in the posterior distribution?

=> 2


# 4E3. 

Using the model definition above, write down the appropriate form of Bayes’ theorem that includes the proper likelihood and priors.


\[
\Pr(\mu, \sigma \mid y_{i}) = \frac{\prod_i \text{Normal}(y_i \mid \mu, \sigma) \text{Normal}(\mu \mid 0, 10) \text{Exponential}(\sigma \mid 1)}{\int \int \prod_i \text{Normal}(y_i \mid \mu, \sigma) \text{Normal}(\mu \mid 0, 10) \text{Exponential}(\sigma \mid 1) \, d\mu \, d\sigma}
\]


# 4E4. 

In the model definition below, which line is the linear model? 

yi ∼ Normal(μ, σ) 

*μi = α + βxi*

α ∼ Normal(0, 10) 

β ∼ Normal(0, 1) 

σ ∼ Exponential(2)


# 4E5. 

In the model definition just above, how many parameters are in the posterior distribution?

=> 3



# 4M1. 
For the model definition below, simulate observed y values from the prior (not the posterior). 

yi ∼ Normal(μ, σ) 
μ ∼ Normal(0, 10) 
σ ∼ Exponential(1)

```{r}

n_values <- 1000
sigma <- rexp(n_values, rate = 1)
mu <- rnorm(n_values, mean = 0, sd = 10)

y <- rnorm(n_values, mu, sigma)

hist(y)
```


# 4M2
Translate the model just above into a quap formula. (and also doing brms formula)

```{r}

# quap: 
flist <- alist( y_i ~ dnorm( mu , sigma ) , 
                mu ~ dnorm( 0 , 10 ) , 
                sigma ~ exp( 1 ) )


# brms model fitting & parameter recovery


library(brms)
formula <- bf(outcome ~ 1)

prior <- c(prior(normal(0,0.1), class = Intercept),
           prior(exponential(1), class = sigma))
  

# can we recover parameters? 
test_data <- tibble(
  outcome = rnorm(100, mean = 2, sd = 2)
                    ) # these mean, and sd should be recovered with a brms model

test_model_100 <- brm(formula = formula, 
                  family = gaussian(), 
                  prior = prior, 
                  data = test_data, 
                  file = "fits/4m2")

plot(test_model_100) # check if chains have mixed well

a <- test_model_100 |> tidybayes::spread_draws(b_Intercept, sigma)

# see if parameters were recovered with 100 data points
hist(a$b_Intercept)
hist(a$sigma)


# now try again with much more data: 
test_data_2 <- tibble(
  outcome = rnorm(10000, mean = 2, sd = 2)
                    )

test_model_10000 <- brm(formula = formula, 
                  family = gaussian(), 
                  prior = prior, 
                  data = test_data_2, 
                  file = "fits/4m2b")

plot(test_model_10000)

a <- test_model_10000 |> spread_draws(b_Intercept, sigma)

hist(a$b_Intercept) # now the posterior is now much more narrow around the sigma and mu that were use to create the data
hist(a$sigma)


```


# 4M3. 
Translate the quap model formula below into a mathematical model definition. 
y ~ dnorm( mu , sigma )

mu <- a + b*x

a ~ dnorm( 0 , 10 )

b ~ dunif( 0 , 1 )

sigma ~ dexp( 1 )


$$
y_i \sim \text{Normal}(\mu, \sigma)
$$
$$
\mu =  \alpha + \beta*x   
$$
$$
\alpha \sim \text{Normal}(0, 10)
$$

$$
\beta \sim \text{Normal}(0, 1)
$$

$$
\sigma \sim \text{Exponential}(1)
$$


# 4M4. 
A sample of students is measured for height each year for 3 years. After the third year, you want to fit a linear regression predicting height using year as a predictor. Write down the mathematical model definition for this regression, using any variable names and priors you choose. Be prepared to defend your choice of priors.


height[i] ~ normmal(mu, sigma)

mu ~ a + b*year

a ~ normal(137, 10)  
(assuming students are children, aged 10, [WHO growth percentiles](https://www.cdc.gov/growthcharts/data/set1clinical/cj41c022.pdf))

b ~ normal(mean = 7, sd = 4) & no value below 0 OR: 

b ~ lognormal(0,1)


sigma ~ exponential(1)


# 4M5
Now suppose I remind you that every student got taller each year. Does this information lead you to change your choice of priors? How?

=> no, the distributions for b do not contain negative numbers

# 4M6
Now suppose I tell you that the variance among heights for students of the same age is never more than 64cm. How does this lead you to revise your priors?


=> variance of 64 => sigma of 8
```{r}
# trying around to inspect the implications of the sigma prior
a <- rexp(1000, rate = 1)

mode_hdi(a)
hist(a)

```

=> not much changes. with exp(1), the value of sigma being above 8 is quite low. Alternatively, we might set a hard limit at 8 for sigma in the prior specification.   


# 4M7

Refit model m4.3 from the chapter, but omit the mean weight xbar this time. Compare the new model’s posterior to that of the original model. In particular, look at the covariance among the parameters. What is different? Then compare the posterior predictions of both models.

```{r}

# get the data
d <- read.csv("https://raw.githubusercontent.com/rmcelreath/rethinking/refs/heads/master/data/Howell1.csv", sep = ";")

d2 <- d |> 
  filter(age >= 18)

d_fit <- d2 |> 
  mutate(mean_weight = mean(weight),
         centered_weight = weight - mean_weight) # adding xbar to the data


# fitting model from textbook
m4.3_original <- brm(height ~ 1 + centered_weight, 
                     family = gaussian, 
                     data = d_fit, 
                     prior = c(prior(normal(178,20), class = Intercept),
                               prior(normal(0, 10), class = b),
                               prior(uniform(0, 50), class = sigma, ub = 50)),
                     iter = 6000, 
                     warmup = 3000, 
                     init = 0,
                     chains = 4, 
                     cores = 4,
                     seed = 4,
                     file = "fits/b04.03")

  

# now with original weight, not centrered
m4.3_original_2 <- brm(height ~ 1 + weight, 
                     family = gaussian, 
                     data = d_fit, 
                     prior = c(prior(normal(178,20), class = Intercept),
                               prior(normal(0, 10), class = b),
                               prior(uniform(0, 50), class = sigma, ub = 50)),
                     iter = 6000, 
                     warmup = 3000, 
                     init = 0,
                     chains = 4, 
                     cores = 4,
                     seed = 4,
                     file = "fits/b04.03b")

# inspect the correlations between the parameters
as_draws_df(m4.3_original)  |>  
  select(b_Intercept:sigma) |>
  cor() |>
  round(digits = 2)
  
as_draws_df(m4.3_original_2) |>
  select(b_Intercept:sigma) |>
  cor() |>
  round(digits = 2) # high correlation between intercept and b
  
  
plot(m4.3_original) 
plot(m4.3_original_2) # intecept from the uncentered is lower
```


# 4M8. 
In the chapter, we used 15 knots with the cherry blossom spline. Increase the number of knots and observe what happens to the resulting spline. Then adjust also the width of the prior on the weights—change the standard deviation of the prior and watch what happens. What do you think the combination of knot number and the prior on the weights controls?


```{r}

# the code for the model is quite long and to do this exercise, we need to play around with values and re-run all of the code. => perfect opportunity for a function. 
cher <- read.csv("https://raw.githubusercontent.com/rmcelreath/rethinking/refs/heads/master/data/cherry_blossoms.csv", sep = ";")


do_splines <- function(n_knots = 15, 
                       width_prior = 10){
    
  # following the code from Solomon Kurz: https://bookdown.org/content/4857/geocentric-models.html#splines.
  
  d2 <-
    cher |> 
    drop_na(doy) # drop incomplete lines
  
  num_knots <- n_knots # this needs to be played around with
  knot_list <- quantile(d2$year, probs = seq(from = 0, to = 1, length.out = num_knots)) # placing the knots at data-informed locations
  
  B <- splines::bs(d2$year,
          knots = knot_list[-c(1, num_knots)], 
          degree = 3, 
          intercept = TRUE) # make the knots
  
  
  d3 <-
    d2 |> 
    mutate(B = B) # add the knots to the data-frame
  
  # some wrangling
  
  n_cols <- n_knots +2
  b <-
    B  |>  
    data.frame() |>  
    set_names(str_c(0, 1:9), 10:n_cols) |>   
    bind_cols(select(d2, year))  |>  
    pivot_longer(-year,
                 names_to = "bias_function",
                 values_to = "bias")
  
  
  
  # fit the model
  local_width_prior <- as.character(width_prior)
  b4.8 <- 
    brm(data = d3,
        family = gaussian,
        doy ~ 1 + B,
        prior = c(prior(normal(100, 10), class = Intercept),
                  prior_string(paste0("normal(0, ",local_width_prior,")"), class = "b"), # these are the widths of the prior, that should be played with for the exercise, therefore, we define this prior as strings
                  prior(exponential(1), class = sigma)),
        iter = 2000, warmup = 1000, chains = 4, cores = 4,
        seed = 4)
  
  
  # visualize the spline
  
  f <- fitted(b4.8)  |>  
    data.frame()  |>  
    bind_cols(d2)  |>  
    
    ggplot(aes(x = year, y = doy, ymin = Q2.5, ymax = Q97.5)) + 
    geom_vline(xintercept = knot_list, color = "white", alpha = 1/2) +
    geom_hline(yintercept = fixef(b4.8)[1, 1], color = "white", linetype = 2) +
    geom_point(color = "#ffb7c5", alpha = 1/2) +
    geom_ribbon(fill = "white", alpha = 2/3) +
    labs(x = "year",
         y = "day in year") +
    theme_bw() +
    theme(panel.background = element_rect(fill = "#4f455c"),
          panel.grid = element_blank())
  
  return(f)
}

# run this on your own machine
# k15_w10 <- do_splines()
# 
# k100_w10 <- do_splines(n_knots = 100)
# 
# k15_w20 <- do_splines(width_prior = 20)
# 
# k15_w1 <- do_splines(width_prior = 1)


```


=> both control in different ways the wigglyness of the line. The n_knows allows for short intervals of the predictor, which makes the line quite adaptive. The width of the weights again gives adaptiveness: the wider the prior of the width of the weights, the more flexible the line. 

# 4H1
The weights listed below were recorded in the !Kung census, but heights were not recorded for these individuals. Provide predicted heights and 89% intervals for each of these individuals. 

=> That is, fill in the table below, using model-based predictions. Individual expected height + 89% interval 

```{r}

dat <- tibble(weight = c(46.95, 43.72, 64.78, 32.59, 54.63)) |> 
  mutate(mean_weight = mean(weight), 
         centered_weight = weight - mean_weight)

d <- read.csv("https://raw.githubusercontent.com/rmcelreath/rethinking/refs/heads/master/data/Howell1.csv", sep = ";") # this is the kung-data from Howell from above, that were used to fit m4.3_original. I thus do not need to refit the model but use the one from M7


# do it once "by hand"

samples_model_original <- tidy_draws(m4.3_original)|> # get draws from the posterior distribution
  mutate(
    pred1 = rnorm(n = n(), 
                       mean =  b_Intercept + b_centered_weight*- 1.584, sd = sigma), 
         pred2 = rnorm(n = n(), mean =  b_Intercept + b_centered_weight*- 4.814, sd = sigma), 
         pred3 = rnorm(n = n(), mean =  b_Intercept + b_centered_weight*16.246, sd = sigma), 
         pred4 = rnorm(n = n(), mean =  b_Intercept + b_centered_weight* -15.944, sd = sigma), 
         pred5 = rnorm(n = n(), mean =  b_Intercept + b_centered_weight*6.096, sd = sigma)
         )

result <- samples_model_original |> 
  select(starts_with("pred")) |> 
  pivot_longer(1:5, names_to = "person", values_to = "height_pred") |> 
  group_by(person) 

result |> 
  ggplot(aes(x = height_pred)) + geom_density() + facet_grid(. ~ person)

result |> 
  summarise(estimate = mode_hdi(height_pred)$y, 
            interval = hdi(height_pred, width = 0.89))

# and again with the predict function from brms
pred_height <-
  predict(m4.3_original,
          newdata = dat, 
          summary = T, 
          probs = c(0.055, 0.945) )

# the two methods lead to sightly different numbers. this is because both version have stochastic processes, visible in the "by hand" version, where we use rnorm(). 


```



# 4H2. 

Select out all the rows in the Howell1 data with ages below 18 years of age. If you do it right, you should end up with a new data frame with 192 rows in it. 

(a) Fit a linear regression to these data, using quap. Present and interpret the estimates. For every 10 units of increase in weight, how much taller does the model predict a child gets? 

*When a child gains 10 kg, it will be 27.2 cm taller.*


(b) Plot the raw data, with height on the vertical axis and weight on the horizontal axis. Superimpose the MAP regression line and 89% interval for the mean. Also superimpose the 89% interval for predicted heights. 

(c) What aspects of the model fit concern you? Describe the kinds of assumptions you would change, if any, to improve the model. You don’t have to write any new code. 

- the lower and upper extremes of weight are overestimated, the middle part is underestimated
- parameter correlation seems to be ok. 
- probably splines would do better, one with a knot at mu-10 and then one at mu+12, or a thrid grade polinomial. 



```{r}

d <- read.csv("https://raw.githubusercontent.com/rmcelreath/rethinking/refs/heads/master/data/Howell1.csv", sep = ";") #load again data
d_kids <- d |> 
  filter(age < 18) |>  # correct
  mutate(mean_weight = mean(weight),
         centered_weight = weight - mean_weight
         )

hist(d_kids$weight)
# a)
h2_kids <- brm(height ~ 1 + centered_weight, 
                     family = gaussian, 
                     data = d_kids, 
                     prior = c(prior(uniform(42, 200), 
                                     class = Intercept, 
                                     lb = 42, 
                                     ub = 200),
                               prior(normal(0, 10), # or a lognormal
                                     class = b),     
                               # mykid1 155 => 40kg , mykid2 145 => 30
                               prior(uniform(0, 50), 
                                     class = sigma, 
                                     ub = 50)),
                     iter = 6000, 
                     warmup = 3000, 
                     init = 0,
                     chains = 4, 
                     cores = 4,
                     seed = 4,
                     file = "fits/b04.03_kids")

# b)
# plotting 
weight_seq <- tibble(centered_weight = seq(from = -20, to = 30, by = 1))

pred_height <-
  predict(h2_kids,
          newdata = weight_seq) |>
  as_tibble() |>
  bind_cols(weight_seq) |> 
  rename(height = Estimate)

mu_summary <- fitted(h2_kids, 
         newdata = weight_seq) |>
  as_tibble() |>
  # let's tackw on the `weight` values from `weight_seq`
  bind_cols(weight_seq)

d_kids |> 
  ggplot(aes(x = centered_weight, y = height)) + 
  
  geom_ribbon(data = pred_height, 
              aes(ymin = Q2.5, ymax = Q97.5), 
              fill = "grey83") + 
  geom_smooth(data = mu_summary,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              fill = "grey70", color = "black", alpha = 1, linewidth = 1/2) +
  geom_point(alpha = 0.2) 



# c)
# look at correlations
as_draws_df(h2_kids)  |>  
  select(b_Intercept:sigma) |>
  cor() |>
  round(digits = 2)

```



# 4H3. 

Suppose a collegue of yours, who works on allometry, glances at the practice problems just above. Your colleague exclaims, “That’s silly. Everyone knows that it’s only the logarithm of body weight that scales with height!” Let’s take your colleague’s advice and see what happens.

## a.1)
 
Model the relationship between height (cm) and the natural logarithm of weight (log-kg). Use the entire Howell1 data frame, all 544 rows, adults and non-adults. Can you interpret the resulting estimates?

=> The text only talks about using the natural log, no mentioning of centralization. First let's illutrate the data and the log-function. The log-scale tells us about "orders of magnitude" (p.98 in rethinking). That is, an increase in the log-scale of weight of one equals to a doubling of the weight.

```{r echo = F}

#|echo: FALSE

d <- read.csv("https://raw.githubusercontent.com/rmcelreath/rethinking/refs/heads/master/data/Howell1.csv", sep = ";") #load again data

d_fit <- d |> 
  mutate(log_weight = log(weight))

line_data <- tibble(x = seq(from = 0.5, to = 101, length.out = 10000), y = log(x))

helping_lines1 <- data.frame(x1 = 0, 
                            y1 = log(1), 
                            x2 = 1, 
                            y2 = log(1))

helping_lines2 <- data.frame(x1 = 0, 
                            y1 = log(10), 
                            x2 = 10, 
                            y2 = log(10))

helping_lines3 <- data.frame(x1 = 0, 
                            y1 = log(100), 
                            x2 = 100, 
                            y2 = log(100))


helping_lines4 <- data.frame(x1 = 1, 
                            y1 = 0, 
                            x2 = 1, 
                            y2 = log(1))

helping_lines5 <- data.frame(x1 = 10, 
                            y1 = 0, 
                            x2 = 10, 
                            y2 = log(10))

helping_lines6 <- data.frame(x1 = 100, 
                            y1 = 0, 
                            x2 = 100, 
                            y2 = log(100))



d_fit |> 
  ggplot(aes(x = weight, y = log_weight)) + 
  geom_point() + 
  geom_line(data = line_data, 
              aes(x = x, y = y)) + 
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), color = "grey", data = helping_lines1, linetype = 2) +
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), color = "grey", data = helping_lines2, linetype = 2) +
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), color = "grey", data = helping_lines3, linetype = 2) +
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), color = "grey", data = helping_lines4, linetype = 2) +
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), color = "grey", data = helping_lines5, linetype = 2) +
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), color = "grey", data = helping_lines6, linetype = 2) + 
  geom_hline(yintercept = 0)  +
  scale_x_continuous(breaks = c(0, 1, 10, 50, 100)) + 
  scale_y_continuous(breaks = c(0, 1, 2, round(log(10),2), 3, 4, round(log(100), 2)))



```

Now fit the model. We adapt the prior to make it give plausable values on the log(weight) scale. The additive part of the formula is: 

mu = a + b * log(weight)

the priors for a and b must therefore be consistent with the log-transformation of the weight to generate average expected heights in cm. `a` indicates the hight of a person that has a log(weight) = 0, which corresponds to a weight of 1: log(1) = : `r log(1)`. 

For the intercept `a`, we use a uniform(20, 100) prior, which corresponds to a height of for a person  with a weight of 1 kg (a person who's `log(weight) = 0`). 

For the slope `b`, we follow along with the normal(0,10). 

```{r}
h3_fit <- brm(height ~ 1 + log_weight, 
                     family = gaussian, 
                     data = d_fit, 
                     prior = c(prior(uniform(0, 200), # we change this to adapt it to the log data
                                     class = Intercept, 
                                     lb = 0, 
                                     ub = 200),
                               prior(normal(0, 10), 
                                     class = b),     
                               prior(uniform(0, 50), 
                                     class = sigma, 
                                     ub = 50)),
                     iter = 6000, 
                     warmup = 3000, 
                     chains = 4, 
                     cores = 4,
                     seed = 4,
                     sample_prior = T,
                     file = "fits/b04.h3_b_log")

summary(h3_fit)

as_draws_df(h3_fit) |> 
  as_tibble()|> 
  select(b_Intercept, b_log_weight, sigma) |>
  cov() |>
  round(digits = 3)

```


=> the intercept `r -23.58` represents the expected height of a person with a log(weight) = 0 kg, that is, the expected height of a person of 1kg. This is if course unrealistic. 

=> the slope `r 47.02` represents the expected increase in height (cm) as a person increases in weight by 1 on the log(weight scale), that is by one order of magnitude in weight. Thus, from 1kg to 10 kg, the model would predict an increase of 47.02 cm of height. From another increase in weight from 10 kg to 100 kg, the model would predict another growth of 47.02 cm. 

However we might want to center the log(weight), since this lead to better model fit in the past. Also, the correlation between log_weight and intercept is quite substanital. We center the log(weight) around it's mean (which is not the same as the mean(weight)). 


## a.2)
```{r}
d_fit <- d|> 
  mutate(log_weight = log(weight),
         mean_log_weight = mean(log_weight), 
         centr_log_weight = log_weight - mean_log_weight)

h3_fit2 <- brm(height ~ 1 + centr_log_weight, 
                     family = gaussian, 
                     data = d_fit, 
                     prior = c(prior(uniform(42, 200), 
                                     class = Intercept, 
                                     lb = 0, 
                                     ub = 200),
                               prior(normal(0, 10), 
                                     class = b),     
                               prior(uniform(0, 50), 
                                     class = sigma, 
                                     ub = 50)),
                     iter = 6000, 
                     warmup = 3000, 
                     chains = 4, 
                     cores = 4,
                     seed = 4,
                     sample_prior = T,
                     file = "fits/b04.h3_b_log_centered")

summary(h3_fit2)

as_draws_df(h3_fit2) |> 
  as_tibble() |> 
  select(b_Intercept, b_centr_log_weight, sigma) |>
  cov() |>
  round(digits = 3)

```

=> the correlation between b_centr_log_weight and intercept is now much smaller. 

=> The intercept of `138.27` shows the height of a person who's log(weight) == mean(log(weight)) = `r `unique(d_fit$mean_log_weight)`. This corresponds to the expected hight of a person with a weight of `exp(unique(d_fit$mean_log_weight))` kg. 

=> the slope `47.01` indicates how much one person is expected to grow when they go up one unit the centered log_weights. This means for example: 

- from increase in weight from centralized_log_weight = 0 (corresponds to 31 kg) to centralized_log_weight = 1 (that is a log_weight of 4.44233 and a weight of exp(log_weight) of `r exp(4.44233)`), we would expect a person to be 47.01 cm higher than the intercept, i.e. have a height of 185.28 cm. 

- if a person is one centralized-log-weight-unit lighter than avearage (this corresponds to a log_weight of 2.44233 and a weight of exp(log_weight) of `r exp(2.44233)`), we would predict this person to be 138.27 - 47.01  = 91.26 cm small.



## b) 
Begin with this plot: plot( height ~ weight , data = Howell1 ). Then use samples from the quadratic approximate posterior of the model in (a) to superimpose on the plot: (1) the predicted mean height as a function of weight, (2) the 97% interval for the mean, and (3) the 97% interval for predicted heights.
```{r}
# plot both centralized and uncentralized data, and alsa plot the predictions on the log-scale and on the normal scale. 

# uncentralized
weight_seq <- tibble(log_weight = c(1, 2, 3, 4, 4.2))

pred_height <-
  predict(h3_fit,
          newdata = weight_seq) |>
  as_tibble() |>
  bind_cols(weight_seq) |> 
  rename(height = Estimate)

mu_summary <- fitted(h3_fit, 
         newdata = weight_seq) |>
  as_tibble() |>
  # let's tackw on the `weight` values from `weight_seq`
  bind_cols(weight_seq)

uc_log <-  d_fit |> 
  ggplot(aes(x = log_weight, y = height)) + # transform back to kg
  
  geom_ribbon(data = pred_height, 
              aes(ymin = Q2.5, ymax = Q97.5), 
              fill = "grey83") + 
  geom_smooth(data = mu_summary,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              fill = "grey70", color = "black", alpha = 1, linewidth = 1/2) +
  geom_point(alpha = 0.2) 

uc_real <- d_fit |> 
  ggplot(aes(x = exp(log_weight), y = height)) + # transform back to kg
  
  geom_ribbon(data = pred_height, 
              aes(ymin = Q2.5, ymax = Q97.5), 
              fill = "grey83") + 
  geom_smooth(data = mu_summary,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              fill = "grey70", color = "black", alpha = 1, linewidth = 1/2) +
  geom_point(alpha = 0.2) 


# centralized
weight_seq <- tibble(centr_log_weight = c(-3, -2, 1, 0, 1))

pred_height <-
  predict(h3_fit2,
          newdata = weight_seq) |>
  as_tibble() |>
  bind_cols(weight_seq) |> 
  rename(height = Estimate)

mu_summary <- fitted(h3_fit2, 
         newdata = weight_seq) |>
  as_tibble() |>
  # let's tackw on the `weight` values from `weight_seq`
  bind_cols(weight_seq)

c_log <-  d_fit |> 
  ggplot(aes(x = centr_log_weight, y = height)) + # transform back to kg
  
  geom_ribbon(data = pred_height, 
              aes(ymin = Q2.5, ymax = Q97.5), 
              fill = "grey83") + 
  geom_smooth(data = mu_summary,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              fill = "grey70", color = "black", alpha = 1, linewidth = 1/2) +
  geom_point(alpha = 0.2) 

c_real <- d_fit |> 
  ggplot(aes(x = exp(centr_log_weight), y = height)) + # transform back to kg
  
  geom_ribbon(data = pred_height, 
              aes(ymin = Q2.5, ymax = Q97.5), 
              fill = "grey83") + 
  geom_smooth(data = mu_summary,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              fill = "grey70", color = "black", alpha = 1, linewidth = 1/2) +
  geom_point(alpha = 0.2) 



cowplot::plot_grid(uc_log, c_log, uc_real, c_real, 
                   nrow = 2, 
                   ncol = 2,
                   labels = c("uncentered, log", "centered, log", "uncentered, real", "centered, real"))

```

# 4H4
Plot the prior predictive distribution for the parabolic polynomial regression model in the chapter. You can modify the code that plots the linear regression prior predictive distribution. Can you modify the prior distributions of α, β1, and β2 so that the prior predictions stay within the biologically reasonable outcome space? That is to say: Do not try to fit the data by hand. But do try to keep the curves consistent with what you know about height and weight, before seeing these exact data.


H6 
